---
title: "Airplane Accident Severity : Group 4"
author: "Amit Shendge, Disha Kacha, Rachel Culbreath, Saniya Shinde"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# You will need to install it (once) from GitHub.
# library(devtools)
# devtools::install_github("physicsland/ezids")
# Then load the package in your R session.
library(ezids)
library(tidyverse)
library(caret)
library(e1071)
```


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
```

```{r}
# 1. Do not provide answers/comments inside code blocks (like here) -- those are notes between coders/self and will be ignored for grading. 
# 2. Make sure your knitr options are set to include all results/code to be graded in the final document.
# 3. All charts/graphs/tables should have appropriate titles/labels/captions. 
# 4. Compose your answers using inline R code instead of using the code-block output as much as you can. 
# 5. Your grade is also determined by the style. Even if you answer everything correctly, but the .html does not look appealing, you will not get full credit. Pay attention to the details that we mentioned in class/homework and in previous sample .Rmd files. For example, how to use #, ##, ###, ..., bold face, italics, inline codes, tables, ..., {results = "asis"}, use of colors in plots/ggplots, and so forth.
```

The project aims to classify the severity of an airplane based on past incidents. The dataset comprises of 2 files : 
Train.csv: [10000 x 12 excluding the headers] contains Training data
Test.csv: [2500 x 11 excluding the headers] contains Test data

Column Descriptions:

Accident_ID : unique id assigned to each row
Accident_Type_Code : type of accident 
Cabin_Temperature : the last recorded temperature before the incident, measured in degrees Fahrenheit
Turbulence_In_gforces : the recorded/estimated turbulence experienced during the accident 
Control_Metric : an estimation of how much control the pilot had during the incident given the factors at play 
Total_Safety_Complaints :  number of complaints from mechanics prior to the accident 
Days_Since_Inspection : how long the plane went without inspection before the incident
Safety_Score : a measure of how safe the plane was deemed to be 
Violations : number of violations that the aircraft received during inspections 
Severity : a description (4 level factor) on the severity of the crash 

```{r reading_csv, include = TRUE}
airplane = read.csv("C:/Users/Saniya Shinde/OneDrive/Desktop/Projects/Semester I/Data Science/Project 2/train.csv")
```

```{r factoring_variables, include  = TRUE}
airplane$Severity <- as.factor(airplane$Severity)
```

```{r unique_severity, include = TRUE}
unique(airplane$Severity)
```

```{r severity_barchart, include = TRUE}
ggplot(airplane, aes(x = Severity, fill = Severity)) + 
  geom_bar(stat = "count", width = 0.8, color = "black") +
  ggtitle("Airplane Severity Distribution") +
  scale_fill_manual(values = c("red", "yellow", "blue", "purple", "orange")) +  # Specify colors
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Severity", y = "Number of Airplanes")
```

We can see from the above plot that most of the airplane crashes were highly fatal and damaging. There is a slight class imabalance with Significant_Damage_And_Fatalities having fewer records.

```{r qqplot_safetyscore, include = TRUE}
qqnorm(airplane$Safety_Score, main="Q-Q Plot for Safety_Score") 
qqline(airplane$Safety_Score)
```
