---
title: "Airplane Accident Severity : Group 4"
author: "Amit Shendge, Disha Kacha, Rachel Culbreath, Saniya Shinde"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# You will need to install it (once) from GitHub.
# library(devtools)
# devtools::install_github("physicsland/ezids")
# Then load the package in your R session.
library(ezids)
library(tidyverse)
library(caret)
library(e1071)
```


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
```

```{r}
# 1. Do not provide answers/comments inside code blocks (like here) -- those are notes between coders/self and will be ignored for grading. 
# 2. Make sure your knitr options are set to include all results/code to be graded in the final document.
# 3. All charts/graphs/tables should have appropriate titles/labels/captions. 
# 4. Compose your answers using inline R code instead of using the code-block output as much as you can. 
# 5. Your grade is also determined by the style. Even if you answer everything correctly, but the .html does not look appealing, you will not get full credit. Pay attention to the details that we mentioned in class/homework and in previous sample .Rmd files. For example, how to use #, ##, ###, ..., bold face, italics, inline codes, tables, ..., {results = "asis"}, use of colors in plots/ggplots, and so forth.
```

The project aims to classify the severity of an airplane based on past incidents. The dataset comprises of 2 files : 
Train.csv: [10000 x 12 excluding the headers] contains Training data
Test.csv: [2500 x 11 excluding the headers] contains Test data

Column Descriptions:

Accident_ID : unique id assigned to each row
Accident_Type_Code : type of accident 
Cabin_Temperature : the last recorded temperature before the incident, measured in degrees Fahrenheit
Turbulence_In_gforces : the recorded/estimated turbulence experienced during the accident 
Control_Metric : an estimation of how much control the pilot had during the incident given the factors at play 
Total_Safety_Complaints :  number of complaints from mechanics prior to the accident 
Days_Since_Inspection : how long the plane went without inspection before the incident
Safety_Score : a measure of how safe the plane was deemed to be 
Violations : number of violations that the aircraft received during inspections 
Severity : a description (4 level factor) on the severity of the crash 

```{r reading_csv, include = TRUE}
airplane = read.csv("C:/Users/Saniya Shinde/OneDrive/Desktop/Projects/Semester I/Data Science/Project 2/train.csv")
```

```{r factoring_variables, include  = TRUE}
airplane$Severity <- as.factor(airplane$Severity)
```

```{r unique_severity, include = TRUE}
unique(airplane$Severity)
```

```{r severity_barchart, include = TRUE}
ggplot(airplane, aes(x = Severity, fill = Severity)) + 
  geom_bar(stat = "count", width = 0.8, color = "black") +
  ggtitle("Airplane Severity Distribution") +
  scale_fill_manual(values = c("red", "yellow", "blue", "purple", "orange")) +  # Specify colors
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Severity", y = "Number of Airplanes")
```

We can see from the above plot that most of the airplane crashes were highly fatal and damaging. There is a slight class imabalance with Significant_Damage_And_Fatalities having fewer records.

```{r qqplot_safetyscore, include = TRUE}
qqnorm(airplane$Safety_Score, main="Q-Q Plot for Safety_Score") 
qqline(airplane$Safety_Score)
```
The Q-Q Plot shows that the data for Safety_Score is normally distributed with the presence of few outliers.

```{r safety_score_severity_boxplot, include  = TRUE}
ggplot(airplane, aes(x = Severity, y = Safety_Score, fill = Severity)) +
  geom_boxplot(outlier.shape = 4, outlier.color = 'green', outlier.size = 3) +
  labs(title = "Boxplot Showing Safety Score Distribution", x = "Severity", y = "Safety Score") +
  theme_minimal() +  # Apply the 'minimal' theme
  scale_fill_manual(values = c("lightblue", "lightyellow", "lightpink", "grey"), name = "Severity") +  # Set boxplot colors
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Title customization
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels by 45 degrees
    axis.title.x = element_text(size = 10, vjust = 0.5)  # X-axis label customization
  ) +
  scale_y_continuous(breaks = seq(0, 100, by = 10))
```
We can see greater variability in the safety score of airplanes having significant_damage_and_fatalities severity. Airplanes having Minor_Damage_And_Injuries and Significant_Damage_And_Fatalities have almost same median safety score. The median of airplanes having Highly_Fatal_And_Damaging severity have lowest safety score.Airplanes having Highly_Fatal_and_Damaging severity have the lowest safety score and  
Airplanes having Significant_Damage_and_Fatalities severity have the highest safety score.

For checking whether the number of violations is a factor that leads to an increased risk for plane crashes based on severity, we are using the analysis of variance (ANOVA) test. The ANOVA test will help us determine if there are significant differences in the mean number of violations across different levels of the severity factor.

Null Hypothesis (H0) : There is no significant difference in the mean number of violations among different severity levels.

Alternative Hypothesis(H1) : There is a significant difference in the mean number of violations among at least two severity levels

```{r ANOVA, include = TRUE}
# Perform ANOVA
anova_result <- aov(Violations ~ Severity, data = airplane)

# Print the ANOVA summary
summary(anova_result)
```

The small p-value (2e-11) indicates that there is strong evidence to reject the null hypothesis.
There is a significant difference in the mean number of violations among at least two severity levels.
The F value of 17.3 is relatively large, further supporting the evidence of a significant difference. Therefore, based on the ANOVA test, it appears that the number of violations is a factor associated with an increased risk for plane crashes based on severity.

```{r post_hoc_tukey_hsd, include  = TRUE}
tukeyviolationAoV <- TukeyHSD(anova_result)
tukeyviolationAoV
```  
The severity levels "Highly_Fatal_And_Damaging" and "Minor_Damage_And_Injuries" are not significantly different in terms of violations. However, "Highly_Fatal_And_Damaging" has significantly higher mean violations compared to the other severity levels, and there are significant differences between "Significant_Damage_And_Fatalities" and "Significant_Damage_And_Serious_Injuries."

```{r correlation_plot, include  = TRUE}
correlation_value <- cor(airplane$Days_Since_Inspection, airplane$Safety_Score)
print(correlation_value)
```

A correlation of -0.685 suggests that there is a meaningful relationship between the number of days since inspection and the safety score. It seems that, as the time since the last inspection increases, there is a tendency for the safety score to decrease.
