---
title: "Airplane Accident Severity : Group 4"
author: "Amit Shendge, Disha Kacha, Rachel Culbreath, Saniya Shinde"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# You will need to install it (once) from GitHub.
# library(devtools)
# devtools::install_github("physicsland/ezids")
# Then load the package in your R session.
library(ezids)
library(tidyverse)
library(caret)
library(e1071)
library(nnet)
```


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
```

```{r}
# 1. Do not provide answers/comments inside code blocks (like here) -- those are notes between coders/self and will be ignored for grading. 
# 2. Make sure your knitr options are set to include all results/code to be graded in the final document.
# 3. All charts/graphs/tables should have appropriate titles/labels/captions. 
# 4. Compose your answers using inline R code instead of using the code-block output as much as you can. 
# 5. Your grade is also determined by the style. Even if you answer everything correctly, but the .html does not look appealing, you will not get full credit. Pay attention to the details that we mentioned in class/homework and in previous sample .Rmd files. For example, how to use #, ##, ###, ..., bold face, italics, inline codes, tables, ..., {results = "asis"}, use of colors in plots/ggplots, and so forth.
```

The project aims to classify the severity of an airplane based on past incidents. The dataset comprises of 2 files : 
Train.csv: [10000 x 12 excluding the headers] contains Training data

Column Descriptions:

Accident_ID : unique id assigned to each row
Accident_Type_Code : type of accident 
Cabin_Temperature : the last recorded temperature before the incident, measured in degrees Fahrenheit
Turbulence_In_gforces : the recorded/estimated turbulence experienced during the accident 
Control_Metric : an estimation of how much control the pilot had during the incident given the factors at play 
Total_Safety_Complaints :  number of complaints from mechanics prior to the accident 
Days_Since_Inspection : how long the plane went without inspection before the incident
Safety_Score : a measure of how safe the plane was deemed to be 
Violations : number of violations that the aircraft received during inspections 
Severity : a description (4 level factor) on the severity of the crash 

```{r reading_csv, include = TRUE}
airplane = read.csv("C:/Users/Saniya Shinde/OneDrive/Desktop/Projects/Semester I/Data Science/Project 2/Airplane Accidents Severity.csv")
```

```{r factoring_variables, include  = TRUE}
airplane$Severity <- as.factor(airplane$Severity)
```

```{r unique_severity, include = TRUE}
unique(airplane$Severity)
```

```{r severity_barchart, include = TRUE}
library(ggplot2)
ggplot(airplane, aes(x = Severity, fill = Severity)) + 
  geom_bar(stat = "count", width = 0.8, color = "black") +
  ggtitle("Airplane Severity Distribution") +
  scale_fill_manual(values = c("red", "yellow", "blue", "purple", "orange")) +  # Specify colors
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size=5)) +
  labs(x = "Severity", y = "Number of Airplanes")
```

We can see from the above plot that most of the airplane crashes were highly fatal and damaging. There is a slight class imabalance with Significant_Damage_And_Fatalities having fewer records.



```{r safety_score_severity_boxplot, include  = TRUE}
ggplot(airplane, aes(x = Severity, y = Safety_Score, fill = Severity)) +
  geom_boxplot(outlier.shape = 4, outlier.color = 'green', outlier.size = 3) +
  labs(title = "Boxplot Showing Safety Score Distribution", x = "Severity", y = "Safety Score") +
  theme_minimal() +  # Apply the 'minimal' theme
  scale_fill_manual(values = c("lightblue", "lightyellow", "lightpink", "grey"), name = "Severity") +  # Set boxplot colors
  theme(
    plot.title = element_text(size = 8, face = "bold", hjust = 0.5),  # Title customization
    axis.text.x = element_text(angle = 45, hjust = 1,size=6),  # Rotate x-axis labels by 45 degrees
    axis.title.x = element_text(size = 10, vjust = 0.5)  # X-axis label customization
  ) +
  scale_y_continuous(breaks = seq(0, 100, by = 20))
```


Null Hypothesis (H0) : There is no significant difference in the mean number of violations among different severity levels.

Alternative Hypothesis(H1) : There is a significant difference in the mean number of violations among at least two severity levels

```{r ANOVA, include = TRUE}
# Perform ANOVA
anova_result <- aov(Violations ~ Severity, data = airplane)

# Print the ANOVA summary
summary(anova_result)
```

The small p-value (3.2e-11) indicates that there is strong evidence to reject the null hypothesis.
There is a significant difference in the mean number of violations among at least two severity levels.
The F value of 17.3 is relatively large, further supporting the evidence of a significant difference. Therefore, based on the ANOVA test, it appears that the number of violations is a factor associated with an increased risk for plane crashes based on severity.

```{r post_hoc_tukey_hsd, include  = TRUE}
tukeyviolationAoV <- TukeyHSD(anova_result)
tukeyviolationAoV
```  
The severity levels "Highly_Fatal_And_Damaging" and "Minor_Damage_And_Injuries" are not significantly different in terms of violations. However, "Highly_Fatal_And_Damaging" has significantly higher mean violations compared to the other severity levels, and there are significant differences between "Significant_Damage_And_Fatalities" and "Significant_Damage_And_Serious_Injuries."

*Accident type 1-7 denotes the following:*
*1.	Locality (Riral, Mixed, Urban)*
*2.	Predominant Development Type (Farms, woods, residential, industrial, pastures)*
*3.	RoadSurfaceCondition (Dry, Wet, Snow, Ice)*
*4.	WeatherCondition-First (Cloudy, clear, raining, snowy, etc)*
*5.	WeatherCondition-Second (A maximum of two weather conditions may be recorded in the crash, such as rain and severe crosswinds. )*
*6.	WeatherContributedtotheCrash (An indication, in the officer’s opinion as to whether or not weather (for example, smoke or hail) was a contributing factor in the crash. *
*a. Yes*
*b. No*
*c. Unknown )*
*7.	AmbientLight (Dusk, Daylight, Darkness, Dawn)*

## Stacked Bar Plot: Violation vs. Accident Type

```{r stacked_bar_plot_violation_accident_type}
library(ggplot2)

# Bar plot of Violations by Accident_Type_Code
ggplot(airplane, aes(x = factor(Accident_Type_Code), fill = factor(Violations))) +
  geom_bar(position = "stack") +
  ggtitle("Bar Plot of Violations by Accident Type Code") +
  xlab("Accident Type Code") +
  ylab("Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
*In the following R code, we are performing a Welch Two Sample t-test using the t.test function on the 'Turbulence_In_gforces' and 'Adverse_Weather_Metric' columns of the 'airplane' dataset. The test assesses if the means of the two numeric variables are significantly different, providing a t-statistic, degrees of freedom, and p-value for interpretation.*
```{r T Test}
t_test_result <- t.test(airplane$Turbulence_In_gforces,airplane$Adverse_Weather_Metric)

t_test_result
```
*Test Statistic (t): The t-value is 31.468. This value represents how many standard errors the difference between the sample means is. In this case, a higher t-value indicates a greater difference between the means.*
*Degrees of Freedom (df): The degrees of freedom are 12004. In the context of the t-test, higher degrees of freedom contribute to a more reliable test.*
*P-value: The p-value is less than 2.2e-16, which is essentially zero. The p-value represents the probability of observing such extreme results (or more extreme) under the assumption that the null hypothesis is true. A very small p-value (less than the common significance level of 0.05) suggests that you can reject the null hypothesis.*
*Null Hypothesis (alternative hypothesis): The null hypothesis is that the true difference in means is equal to 0. The alternative hypothesis is that the true difference in means is not equal to 0.*
*Confidence Interval: The 95% confidence interval for the difference in means is given as (0.1180202, 0.1337002). This interval provides a range of values within which we can be 95% confident that the true difference in means lies.*
*Sample Estimates (Mean of x, Mean of y): The sample mean of 'Turbulence_In_gforces' is 0.3814950, and the sample mean of 'Adverse_Weather_Metric' is 0.2556348. These values represent the average values of the two variables in your samples.*
*With such a small p-value (p-value < 0.05), you would reject the null hypothesis. This suggests that there is a statistically significant difference between the means of 'Turbulence_In_gforces' and 'Adverse_Weather_Metric'. The confidence interval provides a range of plausible values for the true difference in means, and since it does not include 0, it further supports the rejection of the null hypothesis.*

```{r scatter_plot, include  = TRUE}
ggplot(airplane, aes(x = Days_Since_Inspection, y = Safety_Score)) +
  geom_point() +
  labs(title = "Scatter Plot of Days Since Inspection vs Safety Score", x = "Days Since Inspection", y = "Safety Score") +
  theme_minimal()
```

A correlation of -0.685 suggests that there is a meaningful relationship between the number of days since inspection and the safety score. It seems that, as the time since the last inspection increases, there is a tendency for the safety score to decrease.

We will use the logistic model to predict the severity of accidents.

## Logistic Regression Severity 

```{r logistic_regression, include = TRUE}
model <- multinom(Severity ~ Safety_Score + Days_Since_Inspection + Control_Metric + Accident_Type_Code  +Adverse_Weather_Metric , data = airplane)

summary(model)
training_pred <- predict(model, airplane, type="class")

confusionMatrix(training_pred, airplane$Severity)
```

```{r roc_curve, include = TRUE}

model <- multinom(Severity ~ Safety_Score + Days_Since_Inspection + Control_Metric + Accident_Type_Code + Adverse_Weather_Metric , data = airplane)
probs <- predict(model, airplane, type="probs")

# Assuming 'Severity' has three levels: Low, Medium, High
roc_low <- roc(airplane$Severity, probs[, "Highly_Fatal_And_Damaging"])
roc_medium <- roc(airplane$Severity, probs[, "Minor_Damage_And_Injuries"])
roc_high <- roc(airplane$Severity, probs[, "Significant_Damage_And_Fatalities"])
roc_high2 <- roc(airplane$Severity, probs[, "Significant_Damage_And_Serious_Injuries"])

# Plot ROC Curve
plot(roc_low, main="ROC Curves", col="blue")
plot(roc_medium, col="red")
plot(roc_high, col="green")
plot(roc_high2, col="black")
```
