---
title: "US International Flight Data "
author: "Group 4"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# You will need to install it (once) from GitHub.
# library(devtools)
# devtools::install_github("physicsland/ezids")
# Then load the package in your R session.
library(ezids)
```

```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
```

```{r read_csv, include = TRUE}
us_flight_data<-read.csv("C:/Users/Saniya Shinde/OneDrive/Desktop/Projects/Semester I/Data Science/International_Report_Departures.csv")
str(us_flight_data)
```

```{r sort_columns, include = TRUE}
us_flight_data <- us_flight_data[order(us_flight_data$Year,us_flight_data$Month),]
print(us_flight_data)
```

```{r departure_each_month, include = TRUE}
departures_each_month <- aggregate(data_dte ~ Year + Month, data = us_flight_data, FUN = length)
colnames(departures_each_month) <- c("Year","Month","Total_Flights")
print(departures_each_month)
```

```{r departures_by_season, include = TRUE}
season_data <- departures_each_month %>% 
      mutate(Season = case_when(departures_each_month$Month %in% c(12,1,2) ~ 'Winter',                                       departures_each_month$Month %in% c(3,4,5) ~ 'Spring',
                                departures_each_month$Month %in% c(6,7,8) ~ 'Summer',
                                departures_each_month$Month %in% c(9,10,11) ~ 'Fall'))

departure_each_season <- aggregate(Total_Flights ~ Year + Season, data = season_data, FUN = sum)
departure_each_season
```

```{r bar_plot, include = TRUE}
library(ggplot2)

ggplot(departure_each_season, aes(x = Season, y = Total_Flights, fill = Season)) +
  geom_bar(stat = "identity", width = 0.35) +
  labs(x = "Season", y = "Total Flights", title = "Total Flights per Season") +
  scale_fill_manual(values = c("Winter" = "blue", "Spring" = "green", "Summer" = "orange", "Fall" = "red")) +
  theme_minimal()
```
**In the graph provided below, we have curated data by extracting values from both the 'USA gate' and 'Foreign gate' columns while including the most frequently occurring pairs. To illustrate this data, we have chosen a line graph as our visualization method, as depicted in the figure.**
**Our aim in this analysis is to focus on the more common pairings between USA and foreign gates, thereby shedding light on more obvious patterns and relationships in the data.**

```{r occurences line}

# Combine the values from both columns into pairs
pairs <- paste(us_flight_data$usg_apt, us_flight_data$fg_apt)

# Calculate the occurrences of each unique pair
pair_counts <- table(pairs)

# Sort the pairs by occurrence in descending order
sorted_pairs <- names(sort(pair_counts, decreasing = TRUE))

# Select the top 10 pairs
top_10_pairs <- sorted_pairs[1:10]

# Filter the data to include only the top 10 pairs
filtered_data <- data_file_1[pairs %in% top_10_pairs, ]

# Create a line graph for the top 10 pairs
library(ggplot2)
ggplot(data = as.data.frame(table(paste(filtered_data$usg_apt, filtered_data$fg_apt))), 
       aes(x = Var1, y = Freq, group = 1)) +
  geom_line(color = "dodgerblue", size = 1) +
  labs(title = "Top 10 Occurrences for Unique Pairs of USG and FG",
       x = "Pair", y = "Occurrences") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
**In the presented data, the top 10 occurrences of pairs of gates for airlines were analyzed. Notably, Miami emerged as a prominent destination due to its tourist attractions, resulting in high demand for flights. The government has advised both employees and passengers to arrive early at Miami's airport to mitigate crowds and potential delays. Other significant routes in this analysis include Chicago to Toronto and Alaska to Shanghai.**

**This analysis provides valuable insights into travel patterns and can aid in optimizing airport operations and enhancing the passenger experience. The data allows for tracking the highest number of visits based on the airline gate location and destination, aiding in resource allocation and scheduling for more efficient travel management.**

**In our analysis below, we began by converting the columns 'carrier group' and 'Year' into factors since they represent categorical variables. The conversion into factors allows us to treat these variables appropriately in our statistical analysis.**

**To explore and visualize the data further, we employed the 'outlierKD2' function. This function facilitated the generation of box plots, histograms, and quantile-quantile (qq) plots for the aforementioned categorical variables. These graphical representations serve as valuable tools to detect outliers and gain insights into the distribution and characteristics of the data.**

**In this analysis, it's evident that the diagrams, both with and without outliers, appear identical. This similarity arises from the fact that categorical values, such as those in 'carrier group' and 'Year,' are not typically used for this type of graphical representation. Furthermore, it's worth noting that no null values were identified in this dataset. This observation highlights the unique nature of the variables being studied and reinforces the absence of missing data points in our analysis.**

```{r Carrier Group}
library(ezids)
library(ggplot2)
#Convert it into factor

data_file_1$carriergroup <- factor(data_file_1$carriergroup)
data_file_1$Year <- factor(data_file_1$Year)


foreign_out =outlierKD2(data_file_1,fg_wac,rm=TRUE,boxplt=TRUE,histogram = TRUE,qqplt=TRUE)
us_out =outlierKD2(foreign_out,usg_wac,rm=TRUE,boxplt=TRUE,histogram = TRUE,qqplt=TRUE)

complete_data_changes = us_out
```
```{r Carrier Group 1}
foreign_carrier <- subset(complete_data_changes,complete_data_changes$carriergroup==0)

cat("Foreign Carrier are as follows:\n")
head(foreign_carrier,7)

us_carrier  <- subset(complete_data_changes,complete_data_changes$carriergroup==1)

cat("US Carrier are as follows:\n")
head(us_carrier,7)
```

```{r Contingency Table}

contingency_table <- table(complete_data_changes$Year, complete_data_changes$carriergroup)
row_names <- c("2011", "2012", "2013", "2014", "2015", "2016","2017","2018","2019","2020")
rownames(contingency_table) <- row_names

col_names <- c("Foreign","US")
colnames(contingency_table) <- col_names

xkabledply(contingency_table, title="Contingency table")
```

```{r Chi-square}
contingency_table <- table(complete_data_changes$Year, complete_data_changes$carriergroup)

chi_squared_test <- chisq.test(contingency_table)


print(chi_squared_test)
```
```{r flights_year_carriergroup, include = TRUE}
total_flights <- aggregate(Total ~ Year + carriergroup, us_flight_data, sum)
```
```{r bar_chart, include = TRUE}
ggplot(total_flights, aes(x = factor(Year), y = Total, fill = factor(carriergroup))) +
       geom_bar(stat = "identity",position = "dodge") +
      labs(x = "Year", y = "Total Flights", title = "Total Flights by Year and Carrier Group") +
      scale_fill_manual(values = c("0" = "red", "1" = "blue"), 
                       labels = c("0", "1"),
                       name = "Carrier Group") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r bar_chart, include = TRUE}
ggplot(total_flights, aes(x = factor(Year), y = Total, fill = factor(carriergroup))) +
       geom_bar(stat = "identity",position = "dodge") +
      labs(x = "Year", y = "Total Flights", title = "Total Flights by Year and Carrier Group") +
      scale_fill_manual(values = c("0" = "red", "1" = "blue"), 
                labels = c("Foreign Carrier", "Domestic Carrier"),
                       name = "Carrier Group") +
      theme_minimal() 
```
